European or American? A re-examination of the transatlantic influence over the IMF

### 2025-08-01 manuscript/manuscript.qmd
Merge the standalone para into existing ones

At the same time, existing literature raises questions about the consistency and effectiveness of European influence. Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999]. Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015]. While some proposals have called for the consolidation of EU seats to strengthen European influence, critics argue this may entrench the dominance of larger EU states without necessarily enhancing representativeness [@leech2005a; @copelovitch2010a]. Moreover, concerns about Europe’s voting power exceeding its global economic and demographic weight have generated debates about fairness and legitimacy, particularly from developing countries [@woodward2024]. These institutional tensions may place constraints on Europe’s ability to act decisively, though the extent of such limitations remains contested.

@@

Despite Europe’s institutional advantages, scholars remain skeptical of its actual effectiveness in the IMF. Internal divergences—such as France’s growth-oriented stance versus Germany’s fiscal conservatism—combined with fragmented constituencies, often impede coherent action and reduce policy influence [@miescu2014; @dreher2015; @copelovitch2010a]. Moreover, Europe’s disproportionate voting power relative to its population raises legitimacy concerns and is seen as disadvantaging developing countries [@woodward2024; @leech2005a]. While proposals to consolidate EU seats aim to improve efficiency and counterbalance US dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].

>>>>

At the same time, existing literature raises questions about the consistency and effectiveness of European influence. Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999]. Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015; @copelovitch2010a]. While some proposals have called for the consolidation of EU seats to strengthen European influence, critics argue this may entrench the dominance of larger EU states without necessarily enhancing representativeness [@leech2005a; @copelovitch2010a]. Moreover, concerns about Europe’s voting power exceeding its global economic and demographic weight have generated debates about fairness and legitimacy, particularly from developing countries [@woodward2024; @leech2005a]. These institutional tensions may place constraints on Europe’s ability to act decisively, though the extent of such limitations remains contested.

@@
